---
title: "CS422: Association rules"
output: html_notebook
author: "Vijay K. Gurbani, Ph.D., Illinois Institute of Technology"
---

```{r}
library(arules)
library(arulesViz)
library(httpuv)
setwd("~/Desktop/Assignments&Coursework/422/Hw-2-March4/")
rm(list=ls())
```
#### Read the data directly in as **transactions** and inspect them.
```{r}
trans <- read.transactions("mba-2.csv", sep=",")
summary(trans)
```
#### See the first 5 items in the transactions database
```{r}
inspect(trans[1:5])

```
#### Get familiar with the data
```{r}
itemFrequencyPlot(trans, support = 0.1)
image(trans)
```
### Now, let's run Apriori on the dataset.  Note that we only get one rule.  Why?
```{r}
rules <- apriori(trans)
rm(rules)
```
#### We get one rule since our minsup is set too high (0.1).  Let's reduce it.
```{r}
rules <- apriori(trans, parameter = list(support=0.01))
summary(rules)

```
#### Let's inspect the rules, sorted by confidence
```{r}
inspect(head(rules, by="confidence"))
```
#### We can even interactively plot the rules and examine them.
```{r}
plot(rules, engine="htmlwidget")
```
#### You can drill down into rules that have a certain consequent you are looking for as follows:
```{r}
rules.beer <- apriori(trans, parameter=list(supp=0.01),
                 appearance = list(default="lhs", rhs="beer"))
inspect(sort(rules.beer, decreasing = T, by="support"))
```

