---
title: CS 422 Section 01
output: html_notebook
author: Saptarshi Chatterjee (cwid - 20413922)
--- 
### 1 Exercises

> 1.1 Describe in 2-3 sentences what you expect to achieve from this course.

 After completion, I should be conversant with implementation and application of essential data mining concepts and algorithms.I should also be able to contribute to freelance data-science projects available in upwork.com, freelance.com etc.
 
### 2 Practicum problems 

> 1-A: Load the cars dataset and print it.

```{r}
cars
```


> 1-B: Plot the data in the cars dataset. Provide a title for the graph. The title should be “The Cars Dataset”.

```{r}
plot(cars, main="The Cars Dataset")
```
 
 > 1-C: Print a summary of the cars dataset using the R summary() command.
 
```{r}
summary(cars)
```
 
 > 1-D: What is the maximum speed and minimum distance as shown in the summary() command?
 
```{r echo=FALSE}
speed_summary <- summary(cars$speed)
dist_summary <- summary(cars$dist)
paste("Max Speed ", speed_summary[6])
paste("Min distance ", dist_summary[1])
```
 
### 2.2 Problem 2

> 2-A: Read “students.csv” file into a data frame and add a new row consisting of the following data to the existing rows

```{r}
student_data <- read.delim2("~/Desktop/Assignments&Coursework/422/student.csv", sep=";")
new_student <- data.frame(name='Brad Pitt',id=40051,gpa="2.21",grade="C")
new_data <- rbind(student_data, new_student)
new_data
```

