---
title: "CS 422 Section 01"
author: "Saptarshi Chatterjee (cwid - 20413922)"
subtitle: <h5>Solution to Spring 2018 Homework 3<h5>
output:
  html_notebook: default
--- 

### 2.1 Topic: Locality sensitive hashing

```{r}
#Load libraries
library(textreuse)
library(lsa)
options("digits"=3)
rm(list=ls())

files <- list.files("/Users/diesel/Desktop/Assignments&Coursework/422/hw4/Deliverable/hw4.movies", full.names=T)
corpus <- TextReuseCorpus(files, tokenizer = tokenize_ngrams, n = 5,
                          keep_tokens = TRUE)
```
#### (a) Visualizing our dataset as a characteristic matrix, what is the size (rows and columns) of our characteristic matrix? Programmatically compute the rows and columns (i.e., write R commands to compute the rows and columns).

```{r}
columns <- length(corpus)
rows <- c()
for(docno in 1:columns){
  shinglesInThisDoc <- hashes(corpus[docno])[[1]]
  rows <- union(rows, shinglesInThisDoc) 
}

rowcount <- length(unique(unlist(rows),use.names = TRUE ))

sprintf("Number of Columns are %d , and Rows are %d",columns,  rowcount)

```

####(b) Let’s take a look at user ID 20. For this user, answer the following questions by examining the object returned from corpus[[“user20”]], i.e., (i) How many movies has this user rated? (Hint: Use stringr::str_count() method.) (ii) What are the first five shingles (tokens) for this user?

```{r}
d <- corpus[["user20"]]
l <- length(strsplit(d$content, "\n")[[1]])-1
sprintf("Number of movies rated by the user is %d", l)
print("First 5 shingles are --- ")
head(d$tokens, 5)
```

####(c) Let’s find similar users using the brute-force pairwise method first. Using pairwise_candidates() and pairwise_compare()

```{r}
pairs <- pairwise_candidates(pairwise_compare(corpus, jaccard_similarity, progress=interactive()))
pairs
```

####(i) How many pairs of users have a similarity score of at least 0.60? (ii) How many pairs of users have a similarity score of at least 0.50? (iii) How many pairs of users have a similarity score of at least 0.40? (iv) List all the pair of users who have a similarity score of at least 0.40.


```{r}
cat("\npairs of users have a similarity score of at least 0.60 is : ", nrow(pairs[pairs$score*10 > .60, ]))
cat("\npairs of users have a similarity score of at least 0.50 is : ", nrow(pairs[pairs$score*10 > .50, ]))
cat("\npairs of users have a similarity score of at least 0.40 is : ", nrow(pairs[pairs$score*10 > .40, ]))

pairs[pairs$score*10 > .40, ]

```

###2.2 Topic: Content-based recommendation system (2 Points)



#### Take your Illinois Tech ID (minus the ‘A’ prefix) and perform modulo division by 671

```{r}
x <- 20413922 %% 671
paste("user", x)
```

####Once you have constructed the user’s profile, you will randomly choose 10 movies from the movies.csv file
```{r}
setwd("~/Desktop/Assignments&Coursework/422/hw4/ml-latest-small/")
set.seed(100)
movies <- read.csv("movies.csv", header=T, sep=",")
sample <- sample(1:nrow(movies), 10)
moviesample <- movies[sample, ]
moviesample$movieId
```

#### Building a user profile

```{r}
userProfile <- read.csv("/Users/diesel/Desktop/Assignments&Coursework/422/hw4/Deliverable/userprofile.csv", header=T, sep=",")
allGenere <- c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy",
"Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western")

for (row in 1:nrow(userProfile)) {
  generes <- as.list(strsplit(toString(movies[movies$movieId == userProfile[row, ]$X, ]$genres),"[|]")[[1]])
  for (g in generes) {
    userProfile[row, g] = 1
  }
}

for (g in allGenere) {
    print(sum(userProfile[, g]))
    userProfile["avg", ] <- colMeans(userProfile)
}

userProfile <- userProfile[, 1:21]
userProfile
```


#### Creating Movie profile

```{r}
movieProfile <- read.csv("/Users/diesel/Desktop/Assignments&Coursework/422/hw4/Deliverable/movieprofile.csv", header=T, sep=",")
allGenere <- c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime", "Documentary", "Drama", "Fantasy",
"Film-Noir", "Horror", "IMAX", "Musical", "Mystery", "Romance", "Sci-Fi", "Thriller", "War", "Western")

for (row in 1:nrow(movieProfile)) {
  generes <- as.list(strsplit(toString(movies[movies$movieId == movieProfile[row, ]$X, ]$genres),"[|]")[[1]])
  for (g in generes) {
    movieProfile[row, g] = 1
  }
}
movieProfile <- movieProfile[, 1:21]
movieProfile
```


#### Run Cosine similarity

```{r}
options("digits"=5)
for(i in 1:nrow(movieProfile)){
  score <- lsa::cosine(as.numeric(userProfile["avg",-1]), as.numeric(movieProfile[movieProfile$X == movieProfile[i,'X'],-1]))[[1]]

movieId <- movieProfile[movieProfile$X == movieProfile[i,'X'],1]
title <- as.vector(movies[movies$movieId == movieId,"title"])[1]
cat("Movie ", title, ", and similarity score ", score , "\n")
}
```